---
title: "Rmd Exercise Template"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the data

Load the `gapminder` dataset.

```{r chunk name}
gapminder <- read.csv(here::here("data/gapminder5.csv"), stringsAsFactors=FALSE)
```

## If Statement

Use an if() statement to print a suitable message reporting whether there are any records from 2002 in the gapminder dataset. Now do the same for 2012.

```{r}
if (any(gapminder$year == 2002)) {
   print(paste("Record(s) for the year", 2002, "found."))
} else {
  print(paste("No records for year", 2002))
}

if (any(gapminder$year == 2012)) {
   print(paste("Record(s) for the year", 2012, "found."))
} else {
  print(paste("No records for year", 2012))
}
```

## Loop and If Statements

Write a script that finds the mean life expectancy by country for countries whose population is below the mean for the dataset

```{r}
countries = unique(gapminder$country)
mean_pop = mean(gapminder$pop)
for (i in countries) {
    country_mean = mean(gapminder$pop[gapminder$country==i])
    
    if (country_mean < mean_pop) {
        mean_le = mean(gapminder$lifeExp[gapminder$country==i])
        print(paste("Mean Life Expectancy in", i, "is", mean_le))
    }
}
```

Write a script that loops through the `gapminder` data by continent and prints out whether the mean life expectancy is smaller than 50, between 50 and 70, or greater than 70.

```{r}
conts = unique(gapminder$continent)
for (i in conts) {
    mean_le = mean(gapminder$lifeExp[gapminder$continent==i])
    if (mean_le < 50) {
        print(paste("Mean Life Expectancy in", i, "is less than", 50))
    }
    else if (mean_le > 70) {
        print(paste("Mean Life Expectancy in", i, "is greater than", 70))
    }
    else {
        print(paste("Mean Life Expectancy in", i, "is between", 50, "and ", 70))
    }
}
```

## Exercise: Write Functions

Create a function that given a data frame will print the name of each column and the class of data it contains. Use the gapminder dataset.  Hint: Use `mode()` or `class()` to get the class of the data in each column. Remember that `names()` or `colnames()` returns the name of the columns in a dataset.

```{r}
data_frame_info <- function(df) {
  cols <- names(df)
  for (i in cols) {
    print(paste0(i, ": ", mode(df[, i])))
  }
}
data_frame_info(gapminder)
```

Create a function that given a vector will print the mean and the standard deviation of a **vector**, it will optionally also print the median. Hint: include an argument that takes a boolean (`TRUE`/`FALSE`) operator and then include an `if` statement.

```{r}
vec_stats <- function(vec, operator) {
    print(paste("Mean:", mean(vec)))
    print(paste("Standard Deviation:", sd(vec)))
    if (operator) {
        print(paste("Median:", mean(vec)))
    }
}

vector_info(gapminder$lifeExp, operator = F)
vector_info(gapminder$lifeExp, operator = T)
```

## Analyzing the relationship 

Use what you've learned so far to answer the following questions using the `gapminder` dataset. Be sure to include some visualizations!

1. What is the relationship between GDP per capita and life expectancy? Does this relationship change over time? (Hint: Use the natural log of both variables.)

```{r}
plot(log(gapminder$gdpPercap), log(gapminder$lifeExp))
```

2. Does the relationship between GDP per capita and life expectacy vary by continent? Make sure you divide the Americas into North and South America.

```{r}

```

